### 2015-12-14 PrincipiaMetaphysicaAutomationPart1/doc/document/TheoryOfAbstractObjectsV1.tex
some text improvements

  The motivating research questions for the work presented below include:
  \begin{itemize} 
  \item Can functional type theory, despite the problems as pointed by Zalta and Oppenheimer, 
   be utilized to encode the Theory of Abstract Objects when following the embeddings approach?
  \item How elegant and user-friendly is the resulting formalization? In other
   words, to what extend can Isabelle's  user interface be facilitated to hide 
   unpleasant technicalities of the (extended) embedding from the user?
  \item How far can automation be pushed in the approach? How much user interaction can
   be avoided in the formalization of the (first part) of the Principia Metaphysica? 
  \item Can the consistency of the theory be validated with the available automated 
   reasoning tools?
  \item Can the reasoners eventually even contribute some new knowledge? 
  \ldots
  \end{itemize}%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
The encoding of modal functional type theory in functional type theory as explored in 
  previous work \cite{J23,C40} is simple: modal logic formulas are identified with certain functional 
  type theory formulas of predicate type \isa{i{\isasymRightarrow}bool} (abbreviated as \isa{io} below). 
  Possible worlds are explicitly represented by 
  terms of type  \isa{i}. A modal logic \isa{{\isasymphi}} formula holds for a world \isa{w} if and 
  only if the application \isa{{\isasymphi}\ w} evaluates to true. The definition of the propositional modal logic 
  connectives is then straightforward and it simply realizes the standard translation as a set of equations 
  in functional type theory. The approach has been successfully extended for quantifiers. A crucial 
  aspect thereby is that in simple type theory quantifiers can be treated
  as ordinary logical connectives. No extra binding mechanism is needed since the already existing 
  lambda binding mechanism can be elegantly utilized.

  The challenge here is to appropriately 'restrict' this embedding for modal relational type theory.
   \begin{figure}[t]
  \includegraphics[height=5.5cm]{ModalRelationalTypeTheory.png}\includegraphics[height=4.5cm]{ModalRelationalTypeTheory2.png}
  \caption{Grammar of Modal Relational Type Theory. \label{mmrt}
  Note that two kinds of (complex) formulas are introduced: ones that may have encoding subformulas and 
  ones that don’t. The latter are designated as propositional formulas, the former ones simply as formulas. }
  \end{figure}

@@

  is identified and appropriately propagated. The encapsulated expressions (the polymorphic type \isa{{\isacharprime}a} 
  will be instantiated below) realize the actual modeling of the logic embedding analogous 

@@

Some language constructs in the theory of abstract types, e.g. the actuality operator  

@@

  Exemplification yields propositional formulas and never formulas. In the embedding exemplification 

@@

The Principia Metaphysica supports $n$-ary exemplification constructions. For pragmatical 
  reasons we consider here the cases only for $n\leq 3$.%
\ Exe{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \isanewline

@@

\ Exe{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \ \isanewline

@@

  formulas, and their embeddings are straightforward.%
\ not{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \ {\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymnot}\ {\isacharunderscore}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{5}}{\isadigit{8}}{\isacharbrackright}\ {\isadigit{5}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymnot}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline

@@

\ forallBinder{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{binder}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{8}}{\isacharbrackright}\ {\isadigit{9}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymphi}\ x\ {\isasymequiv}\ \isactrlbold {\isasymforall}\ {\isasymphi}{\isachardoublequoteclose}\isanewline

@@

  are validated by our modeling of \isa{\isactrlbold {\isasymbox}}.%

@@

  Metaphysica only for propositional formulas. ... say more about \isa{\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{0}}} ... Their embedding is 

@@

  to \isa{{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}}.
  For pragmatical reasons we restrict ourselves here to the cases where $n\leq 0$. Binder notation is
  introduced for \isa{\isactrlbold {\isasymlambda}} (... unfortuntaley, I don't know yet how this can be achieved as well 
  for \isa{\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharcomma}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{3}}{\isacharcomma}{\isachardot}{\isachardot}{\isachardot}} ...)%

@@

  \isa{\isactrlbold {\isasymiota}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} to \isa{{\isacharparenleft}THE\ x{\isachardot}\ {\isasymphi}\ cw{\isacharparenright}}, that is Isabelle's inbuilt description operator THE 

@@

  Moreover, application of that is allowed only if the body of \isa{{\isasymPhi}}, computed by clause 
  \isa{{\isacharparenleft}{\isasymPhi}\ de{\isacharparenright}}, is a propositional formula. In this case a term is returned and otherwise error 
  reported (for dummy entity \isa{de}). We again introduce binder notation.%

@@

  \isa{ff} and \isa{err}, representing truth, falsity and error. These values can also 
  be noted as \isa{{\isasymtop}}, \isa{{\isasymbottom}} and \isa{{\isacharasterisk}}. We could, of course, also introduce  
  respective logical connectives for the meta-level, but in our applications so far (see below)

@@

  (\isa{{\isasymphi}} is countersatisfiability) and \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v} (\isa{{\isasymphi}} is invalid).%

@@

  be disabled (in the definition of implication), the  we might run into a problem with the definition 
  of \isa{x\ \isactrlbold {\isacharequal}\isactrlsub E\ y} below. Discuss.%

@@

The next two lemmata show that necessitation holds for arbitrary formulas 

@@

  guarding clause. This has been done in previous work. In a simpler, alternative approach we omit 
  this clause here (since all worlds are reachable from any world in an equivalence relation). The 
  following lemmata, which check various conditions for S5,  ensure that we have indeed 

@@

\isamarkupsubsection{Testing for the Correct Propagation of Syntactical Category Information%

@@

>>>>

  The motivating research questions for the work presented below include:
  \begin{itemize} 
  \item Can functional type theory, despite the problems pointed at by Zalta and Oppenheimer, 
   be utilized to encode the Theory of Abstract Objects when following the embeddings approach?
  \item How elegant and user-friendly is the resulting formalization? 
   To what extend can Isabelle's  user interface be facilitated to hide 
   unpleasant technicalities of the (extended) embedding from the user?
  \item How far can automation be pushed in the approach? How much user interaction can
   be avoided in the formalization of the (first part) of the Principia Metaphysica? 
  \item Can the consistency of the theory be validated with the available automated 
   reasoning tools?
  \item Can the reasoners eventually even contribute some new knowledge? 
  \item Suggestions for improvements in Isabelle? Any particular problems detected in the course 
  of the study?
  \ldots
  \end{itemize}%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
The encoding of modal functional type theory in functional type theory as explored in 
  previous work \cite{J23,C40} is simple: modal logic formulas are identified with certain functional 
  type theory formulas of predicate type \isa{i{\isasymRightarrow}bool} (abbreviated as \isa{io} below). 
  Possible worlds are explicitly represented by 
  terms of type  \isa{i}. A modal logic \isa{{\isasymphi}} formula holds for a world \isa{w} if and 
  only if the application \isa{{\isasymphi}\ w} evaluates to true. The definition of the propositional modal logic 
  connectives is then straightforward and it simply realizes the standard translation as a set of equations 
  in functional type theory. The approach has been successfully extended for quantifiers. A crucial 
  aspect thereby is that in simple type theory quantifiers can be treated
  as ordinary logical connectives. No extra binding mechanism is needed since the already existing 
  lambda binding mechanism can be elegantly utilized.

  The challenge here is to appropriately 'restrict' this embedding for modal relational type theory.
   \begin{figure}[t]
  \includegraphics[height=5.5cm]{ModalRelationalTypeTheory.png}\includegraphics[height=4.5cm]{ModalRelationalTypeTheory2.png}
  \caption{Grammar of Modal Relational Type Theory. \label{mmrt}
  Note that two kinds of (complex) formulas are introduced: ones that may have encoding subformulas and 
  ones that do not. The latter are designated as propositional formulas, the former ones simply as formulas. }
  \end{figure}

@@

  is identified and appropriately propagated. The encapsulated expressions  
  realize the actual modeling of the logic embedding analogous 

@@

Some language constructs in the Principia Metaphysica, e.g. the actuality operator  

@@

  Exemplification yields propositional formulas. In the embedding exemplification 

@@

The Principia Metaphysica supports \isa{n}-ary exemplification constructions. For pragmatical 
  reasons we consider here the cases only for \isa{n{\isasymle}{\isadigit{3}}}.%
\ Exe{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\isanewline

@@

\ Exe{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \isanewline

@@

  formulas, and their embeddings are straightforward. In the case of implication the compound formula
  is a propositional formula only of both subformulas are propositional formulas. If at one is a formula
  and the other one a (propositional) formula, then the compound formula is a formula. In all other
  cases an ERR-Formula is returned.%
\ not{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymnot}\ {\isacharunderscore}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{5}}{\isadigit{8}}{\isacharbrackright}\ {\isadigit{5}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymnot}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline

@@

\ forallBinder{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{binder}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{8}}{\isacharbrackright}\ {\isadigit{9}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymphi}\ x\ {\isasymequiv}\ \isactrlbold {\isasymforall}{\isasymphi}{\isachardoublequoteclose}\isanewline

@@

  are validated by our modeling of \isa{\isactrlbold {\isasymbox}}.  \isa{\isactrlbold {\isasymbox}{\isasymphi}} is supported 
  for formulas  and propositional formulas.%

@@

  Metaphysica only over propositional formulas. ... say more about \isa{\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{0}}} ... Their embedding is 

@@

  to \isa{{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}}, etc.
  For pragmatical reasons we restrict ourselves here to the cases where $n\leq 3$. Binder notation is
  introduced for \isa{\isactrlbold {\isasymlambda}} (... unfortuntaley, I don't know yet how binder notation can be 
  achieved also for \isa{\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharcomma}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{3}}{\isacharcomma}{\isachardot}{\isachardot}{\isachardot}} ... need to find out.).%

@@

  \isa{\isactrlbold {\isasymiota}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} to \isa{{\isacharparenleft}THE\ x{\isachardot}\ {\isasymphi}\ cw{\isacharparenright}}, that is Isabelle's inbuilt definite description operator THE 

@@

  Moreover, application of \isa{\isactrlbold {\isasymiota}} to \isa{\isactrlbold {\isasymPhi}} is allowed only if the body of \isa{{\isasymPhi}}, 
  computed by clause \isa{{\isacharparenleft}{\isasymPhi}\ de{\isacharparenright}}, is a propositional formula. In this case a term is returned 
  and otherwise an ERR-term is returned. We again introduce binder notation for \isa{\isactrlbold {\isasymiota}}.%

@@

  \isa{ff} and \isa{err}, representing truth, falsity and error. These values aro also 
  noted as \isa{{\isasymtop}}, \isa{{\isasymbottom}} and \isa{{\isacharasterisk}}. We could, of course, also introduce  
  respective logical connectives for the meta-level, but in our applications (see below)

@@

  (\isa{{\isasymphi}} is countersatisfiability) and \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v} (\isa{{\isasymphi}} is invalid). Actually, so far 
  we only use validity.%

@@

  be disabled (I guess that should then happen in the definition of implication), then we might run 
  into a problem with the definition of \isa{x\ \isactrlbold {\isacharequal}\isactrlsub E\ y} below. Discuss.%

@@

The next two lemmata show that neccessitation holds for arbitrary formulas 

@@

  guarding clause. This has been done in previous work. Our alternative, simpler definition of 
  \isa{{\isasymbox}} above omits
  this clause (since all worlds are reachable from any world in an equivalence relation). The 
  following lemmata, which check various conditions for S5, confirm that we have indeed 

@@

\isamarkupsubsection{Testing the Propagation of Syntactical Category Information%

@@

\ {\isachardoublequoteopen}{\isasymlparr}\isactrlbold {\isasymlambda}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\ \ \ \ \ %
\isamarkupcmt{X is \isa{dio\isactrlsup E}%
